---
title: "Use in Packages"
vignette: >
  %\VignetteIndexEntry{Use in Packages}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
### Table of Contents

2. [Asynchronous Parallel Map](#asynchronous-parallel-map)
3. [Using mirai in a Package](#using-mirai-in-a-package)

### Asynchronous Parallel Map

`mirai_map()` performs asynchronous parallel/distributed map using `mirai`.

This function is similar to `purrr::map()`, but returns a 'mirai_map' object.
It is also more advanced as it allows multiple map over the rows of a dataframe or matrix - and can in fact be used to implement all map variations from that package.

The results of a mirai_map `x` may be collected using `x[]`.
This waits for all asynchronous operations to complete if still in progress.

#### Key advantages:

1. Returns immediately with all evaluations taking place asynchronously. Printing a 'mirai map' object shows the current completion progress.
1. The '.promise' argument allows a promise to registered against each mirai, which can be used to perform side-effects.
1. Returns evaluation errors as 'miraiError' or 'errorValue' as the case may be, rather than causing the entire operation to fail. This allows more efficient recovery from partial failure.
1. Does not rely on a 'chunking' algorithm that attempts to split work into batches according to the number of available daemons, as implemented for example in the `parallel` package. Chunking cannot take into account varying or unpredictable compute times over the indices. It can be optimal to rely on `mirai` for scheduling instead. This is demonstrated in the example below.
```{r}
#| label: chunk
library(mirai)
library(parallel)
cl <- make_cluster(4)
daemons(4)
vec <- c(1, 1, 4, 4, 1, 1, 1, 1)
system.time(mirai_map(vec, Sys.sleep)[])
system.time(parLapply(cl, vec, Sys.sleep))
daemons(0)
```
`.args` is used to specify further constant arguments to `.f` - the 'mean' and 'sd' in the example below:
```{r}
#| label: mmap
with(
  daemons(3, dispatcher = FALSE),
  mirai_map(1:3, rnorm, .args = list(mean = 20, sd = 2))[]
)
```
Use `...` to further specify objects referenced but not defined in `.f` - the 'do' in the anonymous function below:
```{r}
#| label: mmap2
daemons(4)
ml <- mirai_map(
  c(a = 1, b = 2, c = 3),
  function(x) do(x, as.logical(x %% 2)),
  do = nanonext::random
)
ml
ml[]
```
Use of `mirai_map()` requires that `daemons()` have previously been set, and will error if not.

#### Collecting Results

When collecting the results, optionally specify arguments to `[]`:

- `x[.flat]` collects and flattens the results, checking that they are of the same type to avoid coercion.
- `x[.progress]` collects results using a `cli` progress bar, if available, showing completion percentage and ETA, or else a simple text progress indicator of parts completed of the total. If the map operation completes quickly, the `cli` progress bar may not show at all, and this is by design.
- `x[.stop]` collects the results applying early stopping, which stops at the first failure and cancels remaining computations. If the `cli` package is available, it will be used for displaying the error message.

Combinations of the above may be supplied in the fashion of `x[.stop, .progress]`.
```{r}
#| label: mmapadv
#| error: true
mirai_map(list(a = 1, b = "a", c = 3), function(x) exp(x))[.stop]

with(
  daemons(4, dispatcher = FALSE, .compute = "sleep"),
  mirai_map(c(0.1, 0.2, 0.3), Sys.sleep, .compute = "sleep")[.progress, .flat]
)

daemons(0)
```

#### Multiple Map

Multiple map is performed over the **rows** of a dataframe or matrix, as this is most often the desired behaviour.

This allows map over 2 or more arguments by specifying a dataframe.
One of those may be an index value for indexed map.

The function `.f` must take as many arguments as there are columns, either explicitly or via `...`.

```{r}
#| label: mmapmulti
fruit <- c("melon", "grapes", "coconut")

# create a dataframe for indexed map:
df <- data.frame(i = seq_along(fruit), fruit = fruit)

with(
  daemons(3, dispatcher = FALSE),
  mirai_map(df, sprintf, .args = list(fmt = "%d. %s"))[.flat]
)
```
As a dataframe often contains columns of differing type, it is unusual to want to map over the **columns**, however this is possible by simply transforming it beforehand into a list using `as.list()`.

Similarly, the behaviour of `lapply()` or `purrr::map()` on a matrix is the same as that for a vector.
`mirai_map()` on the other hand does take into account the fact that the matrix has dimensions, and maps over its **rows**, consistent with the behaviour for dataframes.
If instead, mapping over the columns is desired, simply take the transpose of the matrix beforehand using `t()`.

[&laquo; Back to ToC](#table-of-contents)

### Using mirai in a Package

mirai as a framework is designed to support completely transparent and inter-operable use within packages.
A core design precept of not relying on global options or environment variables minimises the likelihood of conflict between use by different packages.

There are hence few requirements of package authors, but a few important points to note nevertheless:

1. `daemons()` settings should almost always be left to end-users.
  Consider re-exporting `daemons()` in your package as a convenience.
- Do not include a call to `daemons()` if you use `mirai_map()` or a function that wraps it such as `purrr::map(.parallel = TRUE)`.
  This is important to ensure that there is no unintentional recursive creation of daemons on the same machine.
- Where a `daemons()` call may be appropriate is for async operation using only one dedicated daemon.
  A representative examaple of this usage pattern is `logger::appender_async()`, where the logger package's 'namespace' concept maps directly to mirai's 'compute profile'.

2. The shape and contents of a `status()` call must not be relied upon, as this user interface is subject to change at any time.
- There is a developer interface `nextget()`, for querying values such as 'urls' described in the function documentation.
  Note: only the specifically-documented values are supported interfaces.

3. The functions `unresolved()`, `is_error_value()`, `is_mirai_error()`, and `is_mirai_interrupt()` should be used to test for the relevant state of a mirai or its value.
- The characteristics of how they are currently implemented, e.g. as a logical NA for an 'unresolvedValue', should not be relied upon, as these are subject to change at any time.

4. Testing on CRAN should respect it's 2-core usage limit.
  These limits apply only to tests on CRAN, and more complex tests may be run elsewhere.
- This practically means limiting tests to using one daemon (with `dispatcher = FALSE`) to ensure that only one additional process is used.
- Always reset daemons when done and then allow at least a one-second sleep to ensure all background processes have properly exited.

[&laquo; Back to ToC](#table-of-contents)
