---
title: "2. Async Parallel Map"
vignette: >
  %\VignetteIndexEntry{2. Async Parallel Map}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
### Table of Contents

1. [Introduction](#introduction)
1. [Key Advantages](#key-advantages)
1. [Collecting Results](#collecting-results)
1. [Multiple Map](#multiple-map)

### Introduction

`mirai_map()` performs asynchronous parallel/distributed map using `mirai`.

This function is similar to `purrr::map()`, but returns a 'mirai_map' object.
It is also more advanced as it allows multiple map over the rows of a dataframe or matrix - and can in fact be used to implement all map variations from that package.

The results of a mirai_map `x` may be collected using `x[]`.
This waits for all asynchronous operations to complete if still in progress.

[&laquo; Back to ToC](#table-of-contents)

### Key advantages

1. Returns immediately with all evaluations taking place asynchronously. Printing a 'mirai map' object shows the current completion progress.
1. The '.promise' argument allows a promise to registered against each mirai, which can be used to perform side-effects (such as checkpointing or sending a progress update).
1. Returns evaluation errors as 'miraiError' or 'errorValue' as the case may be, rather than causing the entire operation to fail. This allows more efficient recovery from partial failure.
1. Does not rely on a 'chunking' algorithm that attempts to split work into batches according to the number of available daemons, as implemented for instance in the `parallel` package. 
Chunking cannot take into account varying or unpredictable compute times over the indices, which `mirai` scheduling is designed to deal with optimally. This is demonstrated in the example below.
```{r}
#| label: chunk
library(mirai)
library(parallel)
cl <- make_cluster(4)
daemons(4)
vec <- c(1, 1, 4, 4, 1, 1, 1, 1)
system.time(mirai_map(vec, Sys.sleep)[])
system.time(parLapply(cl, vec, Sys.sleep))
daemons(0)
```
`.args` is used to specify further constant arguments to `.f` - the 'mean' and 'sd' in the example below:
```{r}
#| label: mmap
with(
  daemons(3, dispatcher = FALSE),
  mirai_map(1:3, rnorm, .args = list(mean = 20, sd = 2))[]
)
```
Use `...` to further specify objects referenced but not defined in `.f` - the 'do' in the anonymous function below:
```{r}
#| label: mmap2
daemons(4)
ml <- mirai_map(
  c(a = 1, b = 2, c = 3),
  function(x) do(x, as.logical(x %% 2)),
  do = nanonext::random
)
ml
ml[]
```
Use of `mirai_map()` requires that `daemons()` have previously been set, and will error if not.

[&laquo; Back to ToC](#table-of-contents)

### Collecting Results

When collecting the results, optionally specify arguments to `[]`:

- `x[.flat]` collects and flattens the results, checking that they are of the same type to avoid coercion.
- `x[.progress]` collects results using a `cli` progress bar, if available, showing completion percentage and ETA, or else a simple text progress indicator of parts completed of the total. If the map operation completes quickly, the `cli` progress bar may not show at all, and this is by design.
- `x[.stop]` collects the results applying early stopping, which stops at the first failure and cancels remaining computations. If the `cli` package is available, it will be used for displaying the error message.

Combinations of the above may be supplied in the fashion of `x[.stop, .progress]`.
```{r}
#| label: mmapadv
#| error: true
mirai_map(list(a = 1, b = "a", c = 3), function(x) exp(x))[.stop]

with(
  daemons(4, dispatcher = FALSE, .compute = "sleep"),
  mirai_map(c(0.1, 0.2, 0.3), Sys.sleep)[.progress, .flat]
)
```

[&laquo; Back to ToC](#table-of-contents)

### Multiple Map

Multiple map is performed over the **rows** of a dataframe or matrix, as this is most often the desired behaviour.

> As a dataframe often contains columns of differing type, it is unusual to want to map over the **columns**, however this is possible by simply transforming it beforehand into a list using `as.list()`.

This allows map over 2 or more arguments by specifying a dataframe.
One of those may be an index value for indexed map.

The function `.f` must take as many arguments as there are columns, either explicitly or via `...`.

```{r}
#| label: mmapmulti
fruit <- c("melon", "grapes", "coconut")

# create a dataframe for indexed map:
df <- data.frame(i = seq_along(fruit), fruit = fruit)

with(
  daemons(3, dispatcher = FALSE, .compute = "fruit"),
  mirai_map(df, sprintf, .args = list(fmt = "%d. %s"))[.flat]
)
```

`mirai_map()` maps a matrix over its **rows**, consistent with the behaviour for dataframes
Note that this is different to the behaviour of `lapply()` or `purrr::map()`, which treats a matrix the same as an ordinary vector.

> If instead, mapping over the columns is desired, simply take the transpose of the matrix beforehand using `t()`.

```{r}
#| label: mmapmultimatrix
mat <- matrix(1:4, nrow = 2L, dimnames = list(c("a", "b"), c("y", "z")))
mirai_map(mat, function(x = 10, y = 0, z = 0) x + y + z)[.flat]

daemons(0)
```

[&laquo; Back to ToC](#table-of-contents)
