---
title: "Mirai Map"
vignette: >
  %\VignetteIndexEntry{Mirai Map}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

### 1. Introduction

`mirai_map()` performs asynchronous parallel/distributed map using `mirai`.

This function is similar to `purrr::map()`, but instead of a list of values, returns a 'mirai_map' object, which is a list of mirai.

Incidentally, purrr's `in_parallel()` parallel map capability is based entirely on `mirai_map()`, but does not extend in the same way to async usage, as there you'd wait for completion in all cases.

A `mirai_map()` call returns almost immediately on the other hand.
The results of a mirai_map `x` may be collected using `x[]` or `collect_mirai(x)`, the same as for a mirai.
This waits for all asynchronous operations to complete if still in progress.

Use of `mirai_map()` requires that `daemons()` have previously been set, and will error if this is not the case (rather than launching potentially too many *ephemeral daemons* if the map is over a large number of elements).

### 2. Key Advantages

1. Returns immediately with all evaluations taking place asynchronously. Printing a 'mirai map' object shows the current completion progress.
1. The '.promise' argument allows a promise action to be registered against each iteration. This can be used to perform side-effects when each iteration completes (such as checkpointing or sending a progress update).
1. Returns evaluation errors as 'miraiError' or 'errorValue' as the case may be, rather than causing the entire map to fail. This allows more efficient recovery from partial failure.
1. Does not rely on a 'chunking' algorithm that attempts to split work into batches according to the number of available daemons, as implemented for instance in the `parallel` package. 
Chunking cannot take into account varying or unpredictable compute times over the indices, which `mirai` scheduling is designed to deal with optimally. This is demonstrated in the example below:
```{r}
#| label: chunk
library(mirai)
library(parallel)
cl <- make_cluster(4)
daemons(4)
vec <- c(1, 1, 4, 4, 1, 1, 1, 1)
system.time(mirai_map(vec, Sys.sleep)[])
system.time(parLapply(cl, vec, Sys.sleep))
daemons(0)
```
`.args` is used to specify further constant arguments to `.f` - the `mean` and `sd` in the example below:
```{r}
#| label: mmap
with(
  daemons(3, dispatcher = FALSE),
  mirai_map(1:3, rnorm, .args = list(mean = 20, sd = 2))[]
)
```
Use `...` to further specify objects referenced but not defined in `.f` - the function `do` below:
```{r}
#| label: mmap2
daemons(4)
ml <- mirai_map(
  c(a = 1, b = 2, c = 3),
  function(x) do(x, as.logical(x %% 2)),
  do = nanonext::random
)
ml
ml[]
```

### 3. Collecting Results

When collecting the results, optionally specify arguments to `[]`:

- `x[.flat]` collects and flattens the results, checking that they are of the same type to avoid coercion.
- `x[.progress]` collects results using a `cli` progress bar, if available, showing completion percentage and ETA, or else a simple text progress indicator of parts completed of the total. If the map operation completes quickly, the `cli` progress bar may not show at all, and this is by design.
- `x[.stop]` collects the results applying early stopping, which stops at the first failure and cancels remaining computations. If the `cli` package is available, it will be used for displaying the error message.

Combinations of the above may be supplied in the fashion of `x[.stop, .progress]`.
```{r}
#| label: mmapadv
#| error: true
mirai_map(list(a = 1, b = "a", c = 3), function(x) exp(x))[.stop]

with(
  daemons(4, dispatcher = FALSE, .compute = "sleep"),
  mirai_map(c(0.1, 0.2, 0.3), Sys.sleep)[.progress, .flat]
)
```

### 4. Multiple Map

When a dataframe or matrix is passed as `.x`, multiple map is automatically performed over its **rows**, as this is most often the desired behaviour.

> As a dataframe often contains columns of differing type, it is unusual to want to map over the **columns**, however this is possible by simply transforming it beforehand into a list using `as.list()`.

This allows map over 2 or more arguments by specifying a dataframe.
One of those may be an index value for indexed map.

The function `.f` must take as many arguments as there are columns, either explicitly or via `...`.

```{r}
#| label: mmapmulti
fruit <- c("melon", "grapes", "coconut")

# create a dataframe for indexed map:
df <- data.frame(i = seq_along(fruit), fruit = fruit)

with(
  daemons(3, dispatcher = FALSE, .compute = "fruit"),
  mirai_map(df, sprintf, .args = list(fmt = "%d. %s"))[.flat]
)
```

`mirai_map()` maps a matrix over its **rows**, consistent with the behaviour for dataframes
Note that this is different to the behaviour of `lapply()` or `purrr::map()`, which treats a matrix the same as an ordinary vector.

> If instead, mapping over the columns is desired, simply take the transpose of the matrix beforehand using `t()`.

```{r}
#| label: mmapmultimatrix
mat <- matrix(1:4, nrow = 2L, dimnames = list(c("a", "b"), c("y", "z")))
mirai_map(mat, function(x = 10, y = 0, z = 0) x + y + z)[.flat]

daemons(0)
```

### 5. Nested Maps

In certain cases it may be desirable to perform maps within maps. To do this, the function provided to the outer map needs to include a call to `daemons()` to set daemons used by the inner map.

To prevent accidental proliferation of processes on the same machine, setting local daemons of the form `daemons(6)` from within a function provided to `mirai_map()` will error. As an example, 8 daemons created in the outer map, which each then create 8 in the inner map leads to 64 daemons being created.

It is far more common (and useful) for these to be created across different machines. Creating and launching remote daemons is permitted in all cases.

The above limitation does however prevent a legitimate use pattern: when the outer daemons are launched on remote machines, and you wish to launch inner daemons locally on each of those machines.

The solution in such a case is to use 'remote' daemons which are in fact identical to local daemons.
Instead of a single call to:
```r
daemons(n)
```
Instead, make 2 separate calls to:
```r
â daemons(url = local_url())
launch_local(n)
```
This is permitted from within a mirai map operation.
