---
title: "Communications Backend for R"
vignette: >
  %\VignetteIndexEntry{Communications Backend for R}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

### 1. Mirai Parallel Clusters

mirai provides an alternative communications backend for R.
This functionality was developed to fulfil a request by R Core at R Project Sprint 2023.

A 'miraiCluster' is recognised as one of the official cluster types in R 4.5, and may be created by `parallel::makeCluster(type = "MIRAI")`.
This function calls `make_cluster()`, which may also be used to create a 'miraiCluster' directly.

  + Specify 'n' to launch nodes on the local machine.
  + Specify 'url' for receiving connections from remote nodes.
  + Optionally, specify 'remote' to launch remote daemons using a remote configuration generated by `ssh_config()`, `cluster_config` or `remote_config()`.

Created clusters may be used for any function in the `parallel` base package such as `parallel::clusterApply()` or `parallel::parLapply()`, or the load-balanced versions such as `parallel::parLapplyLB()`.
```{r}
#| label: cluster
library(parallel)
library(mirai)

cl <- makeCluster(6, type = "MIRAI")
cl
parLapply(cl, iris, mean)
```
```{r}
#| label: sleep
#| echo: false
Sys.sleep(1L)
```
`status()` may be called on a 'miraiCluster` to query the number of connected nodes at any time.
```{r}
#| label: cluster2
status(cl)
stopCluster(cl)
```

Making a cluster specifying 'url' without 'remote' causes the shell commands for manual deployment of nodes to be printed to the console.
```{r}
#| label: cluster3
cl <- make_cluster(n = 2, url = host_url())
stop_cluster(cl)
```

### 2. Foreach Support

A 'miraiCluster' may also be registered by [`doParallel`](https://cran.r-project.org/package=doParallel) for use with the [`foreach`](https://cran.r-project.org/package=foreach) package.

Running some parallel examples for the `foreach()` function:

```{r}
#| label: foreach
library(doParallel)
library(foreach)

cl <- makeCluster(6, type = "MIRAI")
registerDoParallel(cl)

# normalize the rows of a matrix
m <- matrix(rnorm(9), 3, 3)
foreach(i = 1:nrow(m), .combine = rbind) %dopar%
  (m[i, ] / mean(m[i, ]))

# simple parallel matrix multiply
a <- matrix(1:16, 4, 4)
b <- t(a)
foreach(b = iterators::iter(b, by='col'), .combine = cbind) %dopar%
  (a %*% b)

stopCluster(cl)
```
